Model-based pretreatment for regression with correlated variables.

### 2014-04-14 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@525 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

We have $I_f=\{I_f^1,\dots,I_f^p \}$ the set of the sets of indices of exogenous covariates (explaining ones $=\boldsymbol{X}_f$) with $\forall j \notin I_r, I_f^j=\emptyset$. Then we have an explicit structure $S=(I_f,I_r,p_f,p_r)$ where $\boldsymbol{p_f}=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each internal regression. Thus we have $p_r=|I_r|$ and $p_f^j=|I_f^j|$ where $|.|$ represents the cardinal of an ensemble.

@@

\subsection{A by-product model: marginal regression with decorrelated covariates}
Now we know $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$ by the structure of sub-regressions, we are able to define a marginal regression model $P(\boldsymbol{Y}|\boldsymbol{X}_f,S)$ based on the reduced set of independent covariates $\boldsymbol{X}_f$ without significant information loss. 
 	\\
Let $\boldsymbol{Y} \in \mathcal{R}^n$ be a response variable we want to explain with $\boldsymbol{X}=[\boldsymbol{X}_f,\boldsymbol{X}_r]$:
	\begin{equation}
			\boldsymbol{Y}_{|\boldsymbol{X}_f,\boldsymbol{X}_r,S}=\boldsymbol{X}_f\boldsymbol{\beta}_f+\boldsymbol{X}_r\boldsymbol{\beta}_r+\boldsymbol{\varepsilon_Y} \textrm{ with } \boldsymbol{\varepsilon_Y} \sim \mathcal{N}(\boldsymbol{0},\sigma^2_Y\boldsymbol{I}_n);	\label{MainR}
		\end{equation}
		where $\boldsymbol{\beta}=(\boldsymbol{\beta}_f,\boldsymbol{\beta}_r) \in  \mathcal{R}^p$ is the vector of the regression coefficients associated respectively to $\boldsymbol{X}_f$ and $\boldsymbol{I}_n$ the identity matrix. 
We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $\boldsymbol{X}_r$ a marginal regression model on $\boldsymbol{Y}$ {\it depending only on uncorrelated covariates $\boldsymbol{X}_f$}:
\begin{eqnarray}
	\boldsymbol{Y}_{|\boldsymbol{X}_f,S}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{Trueexpl} \\
	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*=\boldsymbol{X}\boldsymbol{\beta}^*+\boldsymbol{\varepsilon}_Y^* \textrm{ where }\boldsymbol{\beta}^*=(\boldsymbol{\beta}_f^*,\boldsymbol{\beta}_2^*) \textrm{ and } \hat{\boldsymbol{\beta}}_r^*=\boldsymbol{\beta}_r^*=\boldsymbol{0}
\end{eqnarray}
We note that it simply is a linear regression on some of the original covariates so we only made a pretreatment on the dataset by setting some coefficients to $0$ because of correlations. These $0$ won't be interpreted as independence with the response variable but as redundant information due to correlations between the covariates. It is a variable pre-selection independent of the response $\boldsymbol{Y}$.

@@

Moreover, the uncrossing constraint ($I_1\cap I_2=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.

	The marginal regression model simply is
	\begin{equation}
		\boldsymbol{Y}_{|\boldsymbol{X}_f}= \boldsymbol{X}_f\boldsymbol{\beta}_f^*+ \boldsymbol{\varepsilon}_Y^*
	\end{equation}			
		So we have the \textsc{OLS} unbiased estimator of $\boldsymbol{\beta}^*$: 
		\begin{equation}
			\hat{\boldsymbol{\beta}}_f^* = (\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}\boldsymbol{X}_f'\boldsymbol{Y}  \textrm{ and }\boldsymbol{\hat\beta}_r^* = \boldsymbol{0}
		\end{equation}
		We see in (\ref{Trueexpl}) that it gives an unbiased estimation of $\boldsymbol{Y}$ and $\boldsymbol{\beta^*}$
		but in terms of $\boldsymbol{\beta}$ this estimator is biased:
		\begin{equation}
			\operatorname{E}[\hat{\boldsymbol{\beta}}_f^*|\boldsymbol{X}_f]=\boldsymbol{\beta}_f+\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \textrm{ and }\operatorname{E}[\hat{\boldsymbol{\beta}}_r^*|\boldsymbol{X}_f]=\boldsymbol{0}
		\end{equation}
		with variance:
		\begin{equation}
			\operatorname{Var}[\hat{\boldsymbol{\beta}}_f^*|\boldsymbol{X}_f]= (\sigma^2_Y+\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2 )(\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}  \textrm{ and }\operatorname{Var}[\hat{\boldsymbol{\beta}}_r^*|\boldsymbol{X}_f]= \boldsymbol{0} 
		\end{equation}
		We see that the variance is reduced compared to OLS described in equation (\ref{eqOLS})(no correlations and smaller matrix give better conditioning ) for small values of $\sigma_j$ $i.e.$ strong correlations. So we play on the bias-variance tradeoff, reducing the variance by adding a bias. 		
			%There is no theoretical guarantee that our model is better. It's just a compromise between numerical issues caused by correlations for estimation and selection versus increased variability due to structural hypothesis. Therefore we made some simulations to compare both methods (see the end of this paper).
			 \\
	\subsection{Illustration of the tradeoff conveyed by the pretreatment}		  
	 The Mean Squared Error (\textsc{MSE}) on $\hat{\boldsymbol{\beta}}$ is:
	\begin{eqnarray}
		\textsc{MSE}(\hat{\boldsymbol{\beta}}|\boldsymbol{\beta},\boldsymbol{X})&=&\parallel \operatorname{Bias}\parallel_2^2+\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}})) \\
			&=& 0 + \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}) \textrm{ for OLS, and then for the marginal model:} \\
			&=& \parallel\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1})
	\end{eqnarray}	 
	To better illustrate the bias-variance tradeoff, we look at a simple example with $p=3$ variables. $\boldsymbol{X}_f$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $\boldsymbol{x}_3=\boldsymbol{x}_1+\boldsymbol{x}_2+\boldsymbol{\varepsilon}_3$ where $\boldsymbol{\varepsilon}_3\sim{\mathcal{N}(\boldsymbol{0},\sigma_3^2\boldsymbol{I}_n)}$. We also have $\boldsymbol{\beta}=(1,1,1)$ and $\sigma_Y \in \{1,10\}$  . Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg}'s marginal  model for several values of $\sigma_3$ (strength of the sub-regression) and $n$. Figure \ref{MQE1} shows the theoretical MSE evolution with the strength of the sub-regression:
	\begin{equation}
		1-\mathcal{R}^2=\frac{\operatorname{Var}(\boldsymbol{\varepsilon)_3}}{\operatorname{Var}(\boldsymbol{x}_3)}=\frac{\sigma_3^2}{\sigma_3^2+2}
	\end{equation}

\begin{figure}[h!]
%	\begin{minipage}[l]{.32\linewidth}
%			\includegraphics[width=170px]{figures/MQEn15sigmaY10.png} 
%			\caption{For $n=15$. Dotted: \textsc{Correg}, plain: OLS}\label{MQE1}
%	\end{minipage} \hfill
%	\begin{minipage}[c]{.32\linewidth}
%			\includegraphics[ width=170px]{figures/MQEn100sigmaY10.png} 
%			\caption{For $n=100$. Dotted: \textsc{Correg}, plain: OLS}
%	\end{minipage} \hfill
%   \begin{minipage}[r]{.32\linewidth}
%			\includegraphics[width=170px]{figures/MQEn1000sigmaY10.png} 
%			\caption{For $n=1000$. Dotted: \textsc{Correg}, plain: OLS.} \label{MQE3}
%   \end{minipage} 
	\includegraphics[width=500px]{figures/MQEexplOLS.png}\label{MQE1}
	\caption{MSE of OLS and CorReg (dotted) estimators for varying $(1-R^2)$ of the sub-regression, $n$ and $\sigma_Y$.}
\end{figure} 
It is clear in Figure \ref{MQE1} that the marginal model is more robust than \textsc{OLS} on $X$. And when sub-regression get weaker ($1-\mathcal{R}^2$ tends to 1) it does not explode until extreme values (sub-regression nearly fully explained by the noise). We also see that the error implied by strong correlations shrinks with the rise of $n$. 
We see that $\sigma_Y$ multiplies $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}))=\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{f}))+\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))$ for both models but for the marginal model $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))=0$.
 Thus, as $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}. It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).

\section{Sub-regressions model selection}	
	\subsection{Modeling the uncorrelated covariates: a full generative approach}
%ancienne intro	
Structural equations models like \textsc{SEM} are often used in social sciences and economy where a structure is supposed "by hand" but here we want to find it automatically. Graphical LASSO \cite{friedman2008sparse} offers a method to obtain a structure based on the precision matrix (inverse of the variance-covariance matrix), setting some coefficients of the precision matrix to zero. But the resulting matrix is symmetric and we need an oriented structure for $S$ to avoid cycles. So we developped an MCMC algorithm to find the structure (R package \textsc{CorReg} on CRAN). 
	%ancienne intro

	To be able to compare structures with probabilistic likelihood-based criterions, we need a full generative model on $\boldsymbol{X}$ . We suppose that variables in $\boldsymbol{X_1}$ follow Gaussian mixtures. 
	\begin{equation}
			\forall j \notin I_2 : \boldsymbol{X^j}_{|S} \sim f(\boldsymbol{\theta_j})=\mathcal{GM}(\boldsymbol{\pi_j};\boldsymbol{\mu_j};\boldsymbol{\sigma^2_j}) \textrm{ with } \boldsymbol{\pi_j},\boldsymbol{\mu_j},\boldsymbol{\sigma^2_j} \textrm{ vectors of size } k_j. \label{mixtureX1}
		\end{equation}
		The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust but one can use other laws if needed (exponentially distributed for example). Gaussian case is just a special case ($k_j=1$) of Gaussian mixture so it is included in our hypothesis.

		Variables in $\boldsymbol{X_1}$ are supposed to be independent.
	%Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to use it without impacting the model in other ways. %compute corresponding $\psi$ according to it. %and then improve the walk (will keep a structure only if it is really relevant).%and give it as an input of \textsc{CorReg} and then improve the efficiency of the algorithm (it will find a structure only if it is really relevant).
	We now have a full generative model.
	\subsection{Penalization of the integrated likelihood} \label{compstruct}
  To obtain the marginal regression model we need $S$, the linear structure between the covariates.
  Cross-validation is very time-consuming and thus not friendly with combinatory problematics. Moreover, we need a criterion compatible with structures of different sizes (varying $p_2$) and not related with $Y$ because the structure is inherent to $\boldsymbol{X}$ only. Thus it must be a global criterion. 	
	Because it is about model selection and we are able to provide a full generative model, we decide to follow a Bayesian approach (\cite{raftery1995bayesian}, \cite{andrieu1999joint},\cite{chipman2001practical}).  
	Our full generative generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood of the joint law on $\boldsymbol{X}$ according to the complexity of the structure~\cite{BIChuard}.

	We note $\boldsymbol{\Theta}$ the set of the parameters of the generative model
	\begin{eqnarray}
		P(S|\boldsymbol{X})&\propto& P(\boldsymbol{X}|S)P(S) \label{approxBIC}\\
		-2\log P(\boldsymbol{X}|S)&\approx & BIC=-2\mathcal{L}(\boldsymbol{X},S,\boldsymbol{\Theta})+|\boldsymbol{\Theta}|\log(n)  
	\end{eqnarray}
	But $BIC$ tends to give too complex structures because we test a great range of models. 
	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} law $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$  instead of a simple uniform law on $S$ that is generally used and provides no penalty.
%	 Thus we have :
%		\begin{eqnarray}
%		BIC^*(X|S)&=&BIC(X|S) +\ln(P(S)) \label{Bicstar}
%	\end{eqnarray}		
	It increases penalty on complexity for $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
	this constraint on $\hat{p}_2$ and $\hat{p}_1^j$ is made in the MCMC in the followings.
		We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we will obtain better results when our hypothesis is verified.

	We can also imagine to use other criterions, like the $RIC$ (Risk Inflation Criterion \cite{foster1994risk}) that choose a penalty in $\log p$ instead of $\log n$ and thus gives more parsimonious models when $p$ is larger than $n$ (high dimension) or any other criterion \cite{george1993variable} thought to be better in a given context.

\subsection{MCMC algorithm}
	\subsubsection{The neighbourhood}
	Let's define $\mathcal{S}$ the ensemble of feasible structures (those with $I_1\cap I_2=\emptyset$).
	\\
	For each step, starting from $S \in \mathcal{S}$ we define a neighbourhood:
		\begin{eqnarray}
		\mathcal{V}_{S,j}&=& \{S \}\cup \{ S^{(i,j)} |1\leq i \leq p, i\neq j  \} \\
		\textrm{where }\ \ j &\sim & \mathcal{U}(\{1,\dots,p\}) 
	\end{eqnarray}	
	With $S^{(i,j)}$ defined by the following algorithm :
	\begin{itemize}
		\item if $i \notin I_i^j$ (add): 
			\begin{itemize}
				\item $I_1^j=I_1^j\cup \{i\}$
				\item $I_1^i=\emptyset$ (explicative variables can't depend on others : column-wise relaxation)
				\item $I_1=I_1 \setminus \{j\}$ (dependent variables can't explain others : row-wise relaxation) 
			\end{itemize}			 
		\item else (remove): $I_1^j=I_1^j\setminus \{i\}$
	\end{itemize}

	\smallskip
	At every moment, coherence between $I_1$ and others parts of $S$ can be done by $\forall 1\leq j\leq p :  p_1^j=|I_1^j|$, $I_2=\{j |p_1^j>0 \}$, $p_2= |I_2|$, .

@@

	The walk follows a time-homogeneous Markov Chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we just compute the probabilities when we need them).
	At each step the markov chain moves with probabiliy:
	\begin{eqnarray}
			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=&\sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(-\frac{1}{2}P(\tilde{S}|\boldsymbol{X}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(-\frac{1}{2}P(S_l|\boldsymbol{X}))} \\
	\end{eqnarray}
	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non symétrique mais ne remets  pas en cause l'homogénéité

@@

$n$ & $p_2$ & $P(S)$ &  Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\

>>>>

We have $I_f=\{I_f^1,\dots,I_f^p \}$ the set of the sets of indices of exogenous covariates (explaining ones $=\boldsymbol{X}_f$) with $\forall j \notin I_r, I_f^j=\emptyset$. Then we have an explicit structure $S=(I_f,I_r,p_f,p_r)$ where $p_r=|I_r|$, $\boldsymbol{p}_f=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each internal regression  and $p_f^j=|I_f^j|$. Here $|.|$ represents the cardinal of an ensemble.

@@

\subsection{A by-product model: marginal regression with decorrelated covariates}
Now we know $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$ by the structure of sub-regressions, we are able to define a marginal regression model $P(\boldsymbol{Y}|\boldsymbol{X}_f,S)$ based on the reduced set of independent covariates $\boldsymbol{X}_f$ without significant information loss. 
 	\\
Let $\boldsymbol{Y} \in \mathcal{R}^n$ be a response variable we want to explain with $\boldsymbol{X}=[\boldsymbol{X}_f,\boldsymbol{X}_r]$:
	\begin{equation}
			\boldsymbol{Y}_{|\boldsymbol{X}_f,\boldsymbol{X}_r,S}=\boldsymbol{X}_f\boldsymbol{\beta}_f+\boldsymbol{X}_r\boldsymbol{\beta}_r+\boldsymbol{\varepsilon_Y} \textrm{ with } \boldsymbol{\varepsilon_Y} \sim \mathcal{N}(\boldsymbol{0},\sigma^2_Y\boldsymbol{I}_n);	\label{MainR}
		\end{equation}
		where $\boldsymbol{\beta}=(\boldsymbol{\beta}_f,\boldsymbol{\beta}_r) \in  \mathcal{R}^p$ is the vector of the regression coefficients associated respectively to $\boldsymbol{X}_f$ and $\boldsymbol{I}_n$ the identity matrix. 
We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $\boldsymbol{X}_r$ a marginal regression model on $\boldsymbol{Y}$ {\it depending only on uncorrelated covariates $\boldsymbol{X}_f$}:
\begin{eqnarray}
	\boldsymbol{Y}_{|\boldsymbol{X}_f,S}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{Trueexpl} \\
	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*=\boldsymbol{X}\boldsymbol{\beta}^*+\boldsymbol{\varepsilon}_Y^* \textrm{ where }\boldsymbol{\beta}^*=(\boldsymbol{\beta}_f^*,\boldsymbol{\beta}_2^*) \textrm{ and } \hat{\boldsymbol{\beta}}_r^*=\boldsymbol{\beta}_r^*=\boldsymbol{0}\label{modexpl}
\end{eqnarray}
We note that it simply is a linear regression on some of the original covariates so we only made a pretreatment on the dataset by setting some coefficients to $0$ because of correlations. These $0$ won't be interpreted as independence with the response variable but as redundant information due to correlations between the covariates. It is a variable pre-selection independent of the response $\boldsymbol{Y}$.

@@

Moreover, the uncrossing constraint ($I_f\cap I_r=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.

	For the marginal regression model defined in (\ref{modexpl})
%	\begin{equation}
%		\boldsymbol{Y}_{|\boldsymbol{X}_f}= \boldsymbol{X}_f\boldsymbol{\beta}_f^*+ \boldsymbol{\varepsilon}_Y^*
%	\end{equation}			
%		So 
we have the \textsc{OLS} unbiased estimator of $\boldsymbol{\beta}^*$: 
		\begin{equation}
			\hat{\boldsymbol{\beta}}_f^* = (\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}\boldsymbol{X}_f'\boldsymbol{Y}  \textrm{ and }\boldsymbol{\hat\beta}_r^* = \boldsymbol{0}
		\end{equation}
		We see in (\ref{Trueexpl}) that it gives an unbiased estimation of $\boldsymbol{Y}$ and $\boldsymbol{\beta^*}$
		but in terms of $\boldsymbol{\beta}$ this estimator is biased:
		\begin{equation}
			\operatorname{E}[\hat{\boldsymbol{\beta}}_f^*|\boldsymbol{X}_f]=\boldsymbol{\beta}_f+\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \textrm{ and }\operatorname{E}[\hat{\boldsymbol{\beta}}_r^*|\boldsymbol{X}_f]=\boldsymbol{0}
		\end{equation}
		with variance:
		\begin{equation}
			\operatorname{Var}[\hat{\boldsymbol{\beta}}_f^*|\boldsymbol{X}_f]= (\sigma^2_Y+\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2 )(\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}  \textrm{ and }\operatorname{Var}[\hat{\boldsymbol{\beta}}_r^*|\boldsymbol{X}_f]= \boldsymbol{0} 
		\end{equation}
		We see that the variance is reduced compared to OLS described in equation (\ref{eqOLS})(no correlations and smaller matrix give better conditioning ) for small values of $\sigma_j$ $i.e.$ strong correlations. So we play on the bias-variance tradeoff, reducing the variance by adding a bias. 		
			%There is no theoretical guarantee that our model is better. It's just a compromise between numerical issues caused by correlations for estimation and selection versus increased variability due to structural hypothesis. Therefore we made some simulations to compare both methods (see the end of this paper).
			 \\
	\subsection{Illustration of the tradeoff conveyed by the pretreatment}		  
	 The Mean Squared Error (\textsc{MSE}) on $\hat{\boldsymbol{\beta}}$ is:
	\begin{eqnarray}
		\textsc{MSE}(\hat{\boldsymbol{\beta}}|\boldsymbol{\beta},\boldsymbol{X})&=&\parallel \operatorname{Bias}\parallel_2^2+\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}})) \\
			&=& 0 + \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}) \textrm{ for OLS, and then for the marginal model:} \\
			&=& \parallel\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1})
	\end{eqnarray}	 
	To better illustrate the bias-variance tradeoff, we look at a simple example with $p=3$ variables. $\boldsymbol{X}_f$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $\boldsymbol{x}_3=\boldsymbol{x}_1+\boldsymbol{x}_2+\boldsymbol{\varepsilon}_3$ where $\boldsymbol{\varepsilon}_3\sim{\mathcal{N}(\boldsymbol{0},\sigma_3^2\boldsymbol{I}_n)}$. We also have $\boldsymbol{\beta}=(1,1,1)$ and $\sigma_Y \in \{1,10\}$  . Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg}'s marginal  model for several values of $\sigma_3$ (strength of the sub-regression) and $n$. Figure \ref{MQE1} shows the theoretical MSE evolution with the strength of the sub-regression:
	\begin{equation}
		1-\mathcal{R}^2=\frac{\operatorname{Var}(\boldsymbol{\varepsilon)_3}}{\operatorname{Var}(\boldsymbol{x}_3)}=\frac{\sigma_3^2}{\sigma_3^2+2}
	\end{equation}

\begin{figure}[h!]
%	\begin{minipage}[l]{.32\linewidth}
%			\includegraphics[width=170px]{figures/MQEn15sigmaY10.png} 
%			\caption{For $n=15$. Dotted: \textsc{Correg}, plain: OLS}\label{MQE1}
%	\end{minipage} \hfill
%	\begin{minipage}[c]{.32\linewidth}
%			\includegraphics[ width=170px]{figures/MQEn100sigmaY10.png} 
%			\caption{For $n=100$. Dotted: \textsc{Correg}, plain: OLS}
%	\end{minipage} \hfill
%   \begin{minipage}[r]{.32\linewidth}
%			\includegraphics[width=170px]{figures/MQEn1000sigmaY10.png} 
%			\caption{For $n=1000$. Dotted: \textsc{Correg}, plain: OLS.} \label{MQE3}
%   \end{minipage} 
	\includegraphics[width=500px]{figures/MQEexplOLS.png}\label{MQE1}
	\caption{MSE of OLS and CorReg (dotted) estimators for varying $(1-R^2)$ of the sub-regression, $n$ and $\sigma_Y$.}
\end{figure} 
It is clear in Figure \ref{MQE1} that the marginal model is more robust than \textsc{OLS} on $X$. And when sub-regression get weaker ($1-\mathcal{R}^2$ tends to 1) it remains stable until extreme values (sub-regression nearly fully explained by the noise). We also see that the error implied by strong correlations shrinks with the rise of $n$. 
We see that $\sigma_Y$ multiplies $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}))=\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{f}))+\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))$ for both models but for the marginal model $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))=0$.
 Thus, as $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}. It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).

\section{Sub-regressions model selection}	
Structural equations models like \textsc{SEM} are often used in social sciences and economy where a structure is supposed "by hand" but here we want to find it automatically. Graphical LASSO \cite{friedman2008sparse} offers a method to obtain a structure based on the precision matrix (inverse of the variance-covariance matrix), setting some coefficients of the precision matrix to zero. But the resulting matrix is symmetric and we need an oriented structure for $S$ to avoid cycles.

Cross-validation is very time-consuming and thus not friendly with combinatory problematics. Moreover, we need a criterion compatible with structures of different sizes (varying $p_2$) and not related with $Y$ because the structure is inherent to $\boldsymbol{X}$ only. Thus it must be a global criterion.

We want to find the structure $S$ that maximizes $P(S|\boldsymbol{X})$ and we have:	
	\begin{eqnarray}
	 \label{approxBIC} P(S|\boldsymbol{X})&\propto & P(\boldsymbol{X}|S)P(S)
	=P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)P(\boldsymbol{X}_f|S)P(S)
	\end{eqnarray}

	\subsection{Modeling the uncorrelated covariates: a full generative approach on $P(\boldsymbol{X})$}
	To be able to compare structures with probabilistic likelihood-based criterions, we need a full generative model on $\boldsymbol{X}$. Sub-regressions give $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S) $ but $P(\boldsymbol{X}_f|S)$ is still undefined. We suppose that variables in $\boldsymbol{X}_f$ follow Gaussian mixtures of $k_j \in \mathbf{N}^*$ components: 
	\begin{equation}
			\forall j \notin I_r : \boldsymbol{X}^j_{|S} \sim f(\boldsymbol{\theta}_j)=\mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } k_j. \label{mixtureX1}
		\end{equation}
		The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust but one can use other laws if needed (exponentially distributed for example). Gaussian case is just a special case ($k_j=1$) of Gaussian mixture so it is included in our hypothesis.

		Variables in $\boldsymbol{X}_f$ are in the followings supposed to be independent.
	%Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to use it without impacting the model in other ways. %compute corresponding $\psi$ according to it. %and then improve the walk (will keep a structure only if it is really relevant).%and give it as an input of \textsc{CorReg} and then improve the efficiency of the algorithm (it will find a structure only if it is really relevant).
	We now have a full generative model.

	\subsection{Penalization of the integrated likelihood by $P(S)$} \label{compstruct}

Because it is about model selection and we are able to provide a full generative model, we decide to follow a Bayesian approach (\cite{raftery1995bayesian}, \cite{andrieu1999joint},\cite{chipman2001practical}).

  Our full generative generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood of the joint law on $\boldsymbol{X}$ according to the complexity of the structure~\cite{BIChuard}.

	We note $\boldsymbol{\Theta}$ the set of the parameters of the generative model
	\begin{eqnarray}
		-2\log P(\boldsymbol{X}|S)&\approx & BIC=-2\mathcal{L}(\boldsymbol{X},S,\boldsymbol{\Theta})+|\boldsymbol{\Theta}|\log(n)  
	\end{eqnarray}
	But $BIC$ tends to give too complex structures because we test a great range of models. 
	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} distribution $P(S)=P(I_f | \boldsymbol{p}_f,I_r,p_r)P(\boldsymbol{p}_f|I_r,p_r)P(I_r|p_r)P(p_r)$  instead of a simple uniform law on $S$ that is generally used and provides no penalty.
%	 Thus we have :
%		\begin{eqnarray}
%		BIC^*(X|S)&=&BIC(X|S) +\ln(P(S)) \label{Bicstar}
%	\end{eqnarray}		
	It increases penalty on complexity for $p_r<\frac{p}{2}$ and $p_f^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
	this constraint on $\hat{p}_r$ and $\hat{p}_f^j$ is given in the research algorithm when the Hierarchical Uniform (HU) hypothesis is made instead of Uniform one (U) in numerical experiments (section \ref{sectionsimul} and \ref{sectionrealcase}).
		We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we obtain better results when (HU) this constraint is verified.

%	We can also imagine to use other criterions, like the $RIC$ (Risk Inflation Criterion \cite{foster1994risk}) that choose a penalty in $\log p$ instead of $\log n$ and thus gives more parsimonious models when $p$ is larger than $n$ (high dimension) or any other criterion \cite{george1993variable} thought to be better in a given context. 
%

	%ancienne intro	
	%ancienne intro

\subsection{MCMC algorithm}
	Now we have a comparison criterion, we define an MCMC algorithm to find the structure (R package \textsc{CorReg} on CRAN). 
	\subsubsection{The neighbourhood}
	Let's define $\mathcal{S}$ the ensemble of feasible structures (those with $I_f\cap I_r=\emptyset$).
	\\
	For each step, starting from $S \in \mathcal{S}$ we define a neighbourhood:
		\begin{eqnarray}
		\mathcal{V}_{S,j}&=& \{S \}\cup \{ S^{(i,j)} |1\leq i \leq p, i\neq j  \} \\
		\textrm{where }\ \ j &\sim & \mathcal{U}(\{1,\dots,p\}) 
	\end{eqnarray}	
	With $S^{(i,j)}$ defined by the following algorithm :
	\begin{itemize}
		\item if $i \notin I_f^j$ (add): 
			\begin{itemize}
				\item $I_f^j=I_f^j\cup \{i\}$
				\item $I_f^i=\emptyset$ (explicative variables can't depend on others : column-wise relaxation)
				\item $I_f=I_f \setminus \{j\}$ (dependent variables can't explain others : row-wise relaxation) 
			\end{itemize}			 
		\item else (remove): $I_f^j=I_f^j\setminus \{i\}$
	\end{itemize}

	\smallskip
	At every moment, coherence between $I_f$ and others parts of $S$ can be done by $\forall 1\leq j\leq p :  p_f^j=|I_f^j|$, $I_r=\{j |p_f^j>0 \}$, $p_r= |I_r|$ .

@@

	The walk follows a time-homogeneous Markov Chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we just compute the probabilities when we need them).
	At each step the markov chain moves with probabiliy:
	\begin{eqnarray}
			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=&\sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(-\frac{1}{2}P(\tilde{S}|\boldsymbol{X}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(-\frac{1}{2}P(S_l|\boldsymbol{X}))} 
	\end{eqnarray}
	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non symétrique mais ne remets  pas en cause l'homogénéité

@@

$n$ & $p_r$ & $P(S)$ &  Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_r$ & $\Delta compl$ \\
